---
title: "A11_Noah_McIntire"
author: "Noah McIntire (nem2pq)"
date: "`r Sys.Date()`"
output: html_document
---
# URLs to Shiny Apps
Part 1: https://nem2pq.shinyapps.io/shinyapp_11pt1/
Part 2: http://nem2pq.shinyapps.io/shinyapp_assignment11?_ga=2.45251111.2033156485.1670004182-29570985.1670004182



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(shiny)
```


```{r}
covid <- read.csv('WHO-COVID-19-global-table-data.csv')
covid$mortalityRate <- covid$Deaths...cumulative.total / covid$Cases...cumulative.total
```


```{r}
# make a subset of just 3 countries per region
covid$country <- row.names(covid)
covidSubset <- covid[covid$country %in% c("France", "Germany", "Italy", "Republic of Korea", "Jqpan", "Australia", "Argentina", "Mexico", "Iran (Islamic Republic of)", "Canada", "India", "Indonesia", "Thailand", "South Africa", "Iraq", "Jordan", "Ethiopa", "Kenya", "Zambia"),]
se <- sd(na.omit(covid$mortalityRate))/sqrt(length(na.omit(covid$mortalityRate)))
print(se)
```

```{r}


ui <- fluidPage(
  selectInput("Region", label = "Who Region:", choices = c("Africa", "Americas", "Eastern Mediterranean", "Europe", "South-East Asia", "Western Pacific"), selected = "Africa"),
  plotOutput("mortplot"))


# Calculate SE and add for error bars
server <- function(input, output){
    output$mortplot <- renderPlot({
      covidSubset1 <- covidSubset[covidSubset$Name == input$Region,]
      ggplot(covidSubset1, aes(x=country, y=mortalityRate)) + geom_bar(stat="identity") + ylab('Mortality Rate') +
  xlab('Country') + labs(title="Mortality Rate in Various WHO Regions") + geom_errorbar(aes(ymin = mortalityRate-se, ymax = mortalityRate+se, width = 0.4)) +theme_bw()
    })}

shinyApp(ui = ui, server = server) 
```


# Part 2
https://www.kaggle.com/datasets/whenamancodes/predict-diabities

## Description of the Data Set
This data set from Kaggle, is originally from the National Institute of Diabetes and Digestive and Kidney Diseases. The data set is a potential machine learnig training data set to see if there a predictors that can be used to see if someone has diabetes. This is a selection from a larger database, where all patients selected are of Pima Indian heritage and are at least 21 years of age. Some of the Variables in the data include:
- Pregnancies: Number of pregnancies the individual has had (Continuous)
- Glucose: Count to measure the glucose level in the blood(Continuous)
- Blood Pressure:A numeric value to represent the blood pressure of the individual (Continuous)
- BMI: The body mass index of the individual (Continuous)

From this data set, I have chosen to use Glucose as A and Blood pressure as B, and will explore the relationship between the two with an interactive scatterplot that includes a loess regression.


```{r}
diabetes <- read.csv('diabetes.csv')
ui <- fluidPage(
   sidebarLayout(
    sidebarPanel(
    selectInput("degree", label = h3("Select Degree:"), 
    choices = list( 0, 1, 2), 
    selected = 2),
    sliderInput("span", label = h3("Span adjustment:"),
              min = 0.2, max = 0.9, value = 0.5, step = 0.1)
    ),
  mainPanel(
     plotOutput("diaplot")
    )
  ))

server <- function(input, output){
   
   output$diaplot <- renderPlot({
     
    ggplot(diabetes, aes(x=Glucose, y=BloodPressure)) + geom_point(alpha=0.3) + geom_smooth(method="loess", formula=y ~ poly(x, as.numeric(input$degree)), se=FALSE, span=input$span) + theme_bw()
    })}

shinyApp(ui = ui, server = server) 
```
Based off of the graph produced above, we can see that glucose and blood pressure do not have a relationship of any meaningful value. By changing the degree of the loess line, we can see that it foolows a different path, with 0 creating no line, as it cannot form a relationship with the x axis. The line that seems to fit closest is of degree one, which can either be smoother by increasing the span value or more form fitting by reducing the span value.


